---

title: "PLM Course Project"

output: html_document

---



```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = TRUE)

```



##Overview

This is the course project for Coursera's Practical Machine Learning class.



It uses dataset from http://groupware.les.inf.puc-rio.br/har. Full source:



Velloso, E.; Bulling, A.; Gellersen, H.; Ugulino, W.; Fuks, H. "Qualitative Activity Recognition of Weight Lifting Exercises. Proceedings of 4th International Conference in Cooperation with SIGCHI (Augmented Human '13)". Stuttgart, Germany: ACM SIGCHI, 2013.



##About the Dataset

Train and Validate Data sets were already provided.

The training data is available here:

https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv



The validate data is available here:

https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv



Using devices such as Jawbone Up, Nike FuelBand, and Fitbit it is now possible to collect a large amount of data about personal activity relatively inexpensively. These type of devices are part of the quantified self movement - a group of enthusiasts who take measurements about themselves regularly to improve their health, to find patterns in their behavior, or because they are tech geeks. One thing that people regularly do is quantify how much of a particular activity they do, but they rarely quantify how well they do it.



In this project, we will use data from accelerometers on the belt, forearm, arm, and dumbell of 6 participants. They were asked to perform barbell lifts correctly and incorrectly in 5 different ways.



The goal of the project is to predict the manner in which they did the exercise. This is the "classe" variable in the training set. 


---

title: "PLM Course Project"

output: html_document

---



```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = TRUE)

```



##Overview

This is the course project for Coursera's Practical Machine Learning class.



It uses dataset from http://groupware.les.inf.puc-rio.br/har. Full source:



Velloso, E.; Bulling, A.; Gellersen, H.; Ugulino, W.; Fuks, H. "Qualitative Activity Recognition of Weight Lifting Exercises. Proceedings of 4th International Conference in Cooperation with SIGCHI (Augmented Human '13)". Stuttgart, Germany: ACM SIGCHI, 2013.



##About the Dataset

Train and Validate Data sets were already provided.

The training data is available here:

https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv



The validate data is available here:

https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv



Using devices such as Jawbone Up, Nike FuelBand, and Fitbit it is now possible to collect a large amount of data about personal activity relatively inexpensively. These type of devices are part of the quantified self movement - a group of enthusiasts who take measurements about themselves regularly to improve their health, to find patterns in their behavior, or because they are tech geeks. One thing that people regularly do is quantify how much of a particular activity they do, but they rarely quantify how well they do it.



In this project, we will use data from accelerometers on the belt, forearm, arm, and dumbell of 6 participants. They were asked to perform barbell lifts correctly and incorrectly in 5 different ways.



The goal of the project is to predict the manner in which they did the exercise. This is the "classe" variable in the training set. 

```{r cleaning}

#Delete columns that are purely NA and blank

training <- training[,(colSums(is.na(training) | training=="") == 0)]

validate <- validate[,(colSums(is.na(validate) | validate=="") == 0)]

#Delete irrelevant columns

training <- training[,-(1:5)]

validate <- validate[,-(1:5)]

dim(training)

dim(validate)

```



After cleaning, our final dataset only has 55 columns.



Now that our data is clean, we can now split the data for modelling and prediction.

```{r split}

set.seed(123)

library(caret)

intrain <- createDataPartition(y = training$classe, p = 0.7, list = FALSE)

train <- training[intrain, ]

test <- training[-intrain, ]

```

##Modelling

We will use Random Forest as our model for this project.

```{r train}

library(randomForest)

#Training the model

RFmodel <- randomForest(classe ~ ., data = train, importance = TRUE)

RFmodel

#Prediction

RFpred <- predict(RFmodel, test)

#Confusion Matrix

cm_RF <- confusionMatrix(test$classe, RFpred)

cm_RF

```

Our model has pretty good metrics with 99.75% accuracy.



##Making Predictions

Since our model is pretty good. Let's use it to predict the validate dataset.

```{r validate}

#solves error of mismatch

levels(validate$new_window) <- levels(train$new_window)

predictValidate <- predict(RFmodel, validate[,-55], type="class")

predictValidate

```
